---
title: "PML Course Project"
author: "Viacheslav Shybaiev"
date: "November 23, 2014"
output: html_document
---

# Introduction
This is an Practical Machine Learning course projest.
The assignment objective is to build the learning algorythm to the Human Activity 
Recognitiondata set of 6 participants who were asked to perform barbell lifting 
activities correctly and incorrectly in 5 different ways.

Data set credits:
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

Read more: http://groupware.les.inf.puc-rio.br/har#ixzz3JtzTNxAi

# Processing Data
Reading the raw data from CSV files and building the training model on the *classe* 
factor against the non-empty columns of the training set:
roll_belt, pitch_belt, yaw_belt, gyros_belt_x, 
gyros_belt_y, gyros_belt_z, accel_belt_x, accel_belt_y, 
accel_belt_z, magnet_belt_x, magnet_belt_y, magnet_belt_z, 
roll_arm, pitch_arm, yaw_arm, total_accel_arm, gyros_arm_x, 
gyros_arm_y, gyros_arm_z, accel_arm_x, accel_arm_y, accel_arm_z, 
magnet_arm_x, magnet_arm_y, magnet_arm_z, roll_dumbbell, and yaw_dumbbell.

We're using the Random Forest algorithm for the training model.

```{r, echo=FALSE}
library("caret", lib.loc="/Library/Frameworks/R.framework/Versions/3.1/Resources/library")
library("randomForest", lib.loc="/Library/Frameworks/R.framework/Versions/3.1/Resources/library")
library("lattice", lib.loc="/Library/Frameworks/R.framework/Versions/3.1/Resources/library")
library("ggplot2", lib.loc="/Library/Frameworks/R.framework/Versions/3.1/Resources/library")

```

```{r, echo=TRUE, cache=TRUE}
rawTrain = read.csv ("pml-training.csv")
rawTest = read.csv ("pml-testing.csv")

fit = train (classe ~ roll_belt + pitch_belt + yaw_belt + gyros_belt_x + 
                     gyros_belt_y + gyros_belt_z + accel_belt_x + accel_belt_y + 
                     accel_belt_z + magnet_belt_x + magnet_belt_y + magnet_belt_z + 
                     roll_arm + pitch_arm + yaw_arm + total_accel_arm + gyros_arm_x + 
                     gyros_arm_y + gyros_arm_z + accel_arm_x + accel_arm_y + accel_arm_z + 
                     magnet_arm_x + magnet_arm_y + magnet_arm_z + roll_dumbbell + 
                     yaw_dumbbell, method = "rf", data=rawTrain)

```

## Overview of the model
```{r, echo=TRUE}
print(fit)
```

The test set contsist of 20 test cases which we're going to classify on the 
*classe* factor.

Call predict function from the caret package:
```{r, echo=TRUE}
predict(fit, rawTest)
```

As seen above, there are 20 classifications for each of the test cases.

## Estimate of sample error

We're using plot of the final model to estimate the sample error:

```{r, echo=FALSE}
plot(fit$finalModel, compress = TRUE)
```

## Variable Importance
Further model optimisation could be done using the variable importance order:
```{r, echo=TRUE}
varImp(fit)
```

# Summary
In this assignment we performed very basic machine learning excersise building 
the training model and using the Ranfom Forest algorithm.

